---
output: 
  github_document:
    html_preview: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# semiBRM: Semiparametric Binary Response Model 

This R package `semiBRM` offers an implementation of single-index semiparametric binary response models introduced in [Klein and Spady (1993)](https://doi.org/10.2307/2951556). 

This is built upon `Rcpp` along with `OpenMP`, which parallelizes computation of nonparametric conditional expectations over given data points. This helps improve computation efficiency enormously in parameter estimation, which is one of the main features of this package.

This is still in development. The ultimate goal of it is to help researchers employ [Klein and Spady (1993)](https://doi.org/10.2307/2951556) as easy as running Probit or Logistic regressions in R.

## Identifiable Parameters in Semiparametric Models
In a single-index semiparametric approach, not all parameters are identifiable. First of all, intercept cannot be estimated. Secondly, the coefficients can be estimated as ratios to a "basis" coefficient. This package sets the coefficient of the first explanatory variable as the basis so that its coefficient is normalized to 1, and those of the rest of the variables are estimated conformably to it. 

In the context of binary response models, or nonlinear models more broadly, this is not problematic at all, as the coefficients themselves have little meaning. As the matter of fact, the identifiable set of parameters in this approach can correctly estimate the conditional probability of Y being 1 given X, which would be the ultimate quantity of interest in most of the cases.

## Installation
The development version can be installed from GitHub as follows:

``` r
library(devtools)
install_github(repo="hk599/semiBRM")
```

### This version was built with:

* R 4.0.4
* rtools40 
* maxLik 1.4-6
* matrixStats 0.58.0
* Rcpp 1.0.6
* RcppArmadillo 0.10.2.2.0




## Example

### Implementation
```{r example, message=FALSE, warning=FALSE}
library(semiBRM)

# data generating process
N <- 1500L
X1 <- rnorm(N)
X2 <- (X1 + 2*rnorm(N))/sqrt(5) + 1
X3 <- rnorm(N)^2/sqrt(2)
X <- cbind(X1, X2, X3)
beta <- c(2, 2, -1, -1) # this is an original set of coefficients
V <- as.vector(cbind(X, 1)%*%beta)
Y <- ifelse(V >= rnorm(N), 1L, 0L)


# parameter estimation
coefs_true <- c(1, -.5) # this is an identifiable set of coefficients

## semiparametric approach
qmle <- semiBRM(X = X, Y = Y, control = list(iterlim=50))
coefs_semi <- qmle$estimate

## parametric approach (probit)
probit <- glm(Y ~ X, family = binomial(link = "probit"))
coefs_probit <- probit$coefficients[-1L][-1L]/probit$coefficients[-1L][1L]


# conditional probability at the means
Xbar <- colMeans(X)
Prob_0 <- pnorm( as.vector(c(Xbar, 1)%*%beta) ) # true conditional probability at the means

## semiparametric approach
Vhat <- X[,1L] + as.vector(X[,-1L]%*%coefs_semi)
Vhat_bar <- Xbar[1L] + as.vector(Xbar[-1L]%*%coefs_semi)
h <- sd(Vhat)*N^(-1/5)
Prob_semi <- GaussianKerNonpar(Y, Vhat, Vhat_bar, h)

## parametric approach
Prob_prob <- pnorm(as.vector(probit$coefficients%*%c(1, Xbar)))
```

### Results
```{r results, cache = TRUE, message=FALSE, warning=FALSE, echo=FALSE}
# formatted outputting
{
cat(sprintf("        %7s\t%7s\t%7s", "true", "Semi", "Probit"), fill = TRUE)
cat(sprintf("parm 1: %7.4f\t%7.4f\t%7.4f", coefs_true[1L], coefs_semi[1L], coefs_probit[1L]), fill = TRUE)
cat(sprintf("parm 2: %7.4f\t%7.4f\t%7.4f", coefs_true[2L], coefs_semi[2L], coefs_probit[2L]), fill = TRUE)
cat(sprintf("prob. : %7.4f\t%7.4f\t%7.4f", Prob_0, Prob_semi, Prob_prob), fill = TRUE)
}
```

## References
Klein, R. W., & Spady, R. H. (1993). An Efficient Semiparametric Estimator for Binary Response Models. _Econometrica_, 61(2), 387-421. _[https://doi.org/10.2307/2951556](https://doi.org/10.2307/2951556)_.



